---
title: "Visionary Final"
author: "Justin DePue, John Rooney, and Tony Jiang"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## library packages

```{r warning=FALSE, message=FALSE}

library(tseries)
library(lubridate)
library(here)
library(tidyverse)
library(outliers)
library(ggplot2)
library(forecast)
library(kableExtra)
library(readxl)

```

## load the data

```{r warning=FALSE}

# automatically set the working directory as personal local path to R project
# so the file path in read.csv can work for everyone
setwd(here())

# load the data
electricity_prices.df <- read.csv("./Data/Average_retail_price_of_electricity.csv", header = TRUE, skip=4)

nc_electricity.df<-electricity_prices.df[228,4:(ncol(electricity_prices.df))] %>%
  pivot_longer(cols=everything(), names_to = 'my_date', values_to = 'price_per_kWh')

# examine whether there is missing value or not
summary(nc_electricity.df)
# there is no missing value

nc_electricity.df$price_per_kWh<-as.numeric(nc_electricity.df$price_per_kWh)
  

# Import Natural Gas data

natural_gas.df <- read.csv("./Data/NC_NaturalGas.csv", header = TRUE, skip=2,col.names = c("year", "price"))

#Check for missing data. There is an extra row with an NA at the end of the data, so I removed it with na.omit

summary(na.omit(natural_gas.df))
str(natural_gas.df)

#create timeseries for natural gas

ts_NG<-ts(na.omit(natural_gas.df[,2]), start=c(1989,1), frequency=12)

# plot raw time series
plot(ts_NG, col="blue", ylab="$/Mcf", xlab="Date", main="NC Residential Natural Gas Cost")

```
## standardize the unit of natural gas data to make it comparable with electricity
```{r, warning=FALSE}
# convert natural gas data from $/Mcf to $/kWh based on 80% heating efficiency
# $/kWh = [($/mcf/1.037)/293.07107]/0.9 = $/mcf/273.52


conversion <- 0.8*293.07107*1.037/100


natural_gas.df$kwh_equiv<-((natural_gas.df$price)/conversion)


natural_gas.df<-na.omit(natural_gas.df)

ts_gas_equiv<-ts(natural_gas.df[,3], start=c(1989,1), frequency=12)

ts_gas_equiv<-window(ts_gas_equiv, start=c(2001, 1))

#plot gas TS in kw/hr equiv
autoplot(ts_gas_equiv) +
  ylab("Price (cents/kWh") + 
  ggtitle("Natural Gas Price (cents/kWh)")


# coerce electricity to a time series object

ts_electricity<-ts(rev(nc_electricity.df[,2]), start=c(2001,1), frequency=12)

plot(ts_electricity, col="red", ylab="$/kWh", xlab="Date", main="NC Residential Electricity Cost")

#plot both electric & gas TS together
ts.plot(ts_gas_equiv, ts_electricity, gpars = list(col = c("red", "blue")), 
        xlab="Date", ylab="cents/kWh", 
        main="Comparison of Natural Gas and Electricity Costs")
legend("topleft", bty="n", lty=c(1,1), col=c("red","blue"),
       legend=c(" Natural Gas ", " Electricity "))

#Create Dataframe with both Gas & Electricity prices...
#create new date column
new_date<-seq(as.Date("2001/1/1"), by = "month", length.out = nrow(ts_electricity))

#bind date, gas, & electricity
nc_energy.df<-cbind.data.frame("Month_Date"=new_date, 
                               "electricity"=nc_electricity.df$price_per_kWh, 
                               "gas_equiv"=ts_gas_equiv)

#Subtract Gas price from Electricity price
nc_energy.df$cost_diff<-(nc_energy.df$electricity - nc_energy.df$gas_equiv)

```

## add covariates
```{r}
# create an indicator variable for Ukraine War for gas

# create a column with row index
natural_gas.df <- rownames_to_column(natural_gas.df, var = "index")

natural_gas.df$index = as.numeric(natural_gas.df$index)

natural_gas.df$UKRWAR <- ifelse(natural_gas.df$index >= 399, 1, 0)

natural_gas.df <- natural_gas.df[, c(-1, -6)]

ts_gas_equiv<-ts(natural_gas.df[,c(3,4)], start=c(1989,1), frequency=12) %>% 
  window(start = c(2001, 1))

# create an indicator variable for Ukraine War for electricity
nc_electricity.df <- rownames_to_column(nc_electricity.df, var = "index")

nc_electricity.df$index <- as.numeric(nc_electricity.df$index)

nc_electricity.df$UKRWAR <- ifelse(nc_electricity.df$index >= 255, 1, 0)

nc_electricity.df <- nc_electricity.df[, -1]

# coerce electricity to a time series object

ts_electricity<-ts(rev(nc_electricity.df[,2:3]), start=c(2001,1), frequency=12)


# import temperature data
temperautre <- read_xlsx("./Data/Raleigh Temperature.xlsx") %>% 
  gather(key = "Month", value = "temperature", Jan, Feb, Mar, Apr, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec)

temperautre <- temperautre[order(temperautre$Year), ] 

ts_temperature <- temperautre[,3] %>% 
  na.omit() %>% 
  ts(start = c(1995, 1), frequency = 12) %>% 
  window(start = c(2001, 1), end = c(2023, 1))

# create a data frame for covariates for electricity (2001-2022)
# we need to do this because electricity and natural gas 
# have different fourier terms
fourier_train_e <- fourier(window(ts_electricity[, 2],
                                end = c(2022, 1)),
                         K = 6)

covariates_train_e <- ts_electricity[, 1] %>% 
  window(end = c(2022, 1)) %>% 
  cbind(window(ts_temperature, end = c(2022,1)),
        fourier_train_e)


# create covariates data frame for electricity (2001-2023)
fourier_full_e <- fourier(window(ts_electricity[, 2],
                                end = c(2023, 1)),
                         K = 6)

covariates_full_e <- cbind(ts_electricity[, 1], ts_temperature, fourier_full_e)

# create covariates data frame for natural gas (2001-2022)
fourier_train_gas <- fourier(window(ts_gas_equiv[, 1],
                                end = c(2022, 1)),
                         K = 6)

covariates_train_gas <- ts_gas_equiv[, 2] %>% 
  window(end = c(2022, 1)) %>% 
  cbind(window(ts_temperature, end = c(2022,1)
               ),
        fourier_train_gas)


# create covariates data frame for electricity (2001-2023)
fourier_full_gas <- fourier(window(ts_gas_equiv[, 1],
                                end = c(2023, 1)),
                         K = 6)

covariates_full_gas <- cbind(ts_gas_equiv[, 2], ts_temperature, fourier_full_gas)
  
```

# Decompose TS
```{r}
#Decompose TS

decomp_elec<-decompose(ts_electricity[,2], type="additive")
decomp_gas<-decompose(ts_gas_equiv[,1], type="multiplicative")

ts_gas_equiv[,1] %>% 
  window(start = c(2013,1), end = c(2023,1)) %>%
  stl(s.window = "periodic") %>% 
  plot()

plot(decomp_elec)

#remove seasonality
ts_elec_ns<-(ts_electricity[,2] - decomp_elec$seasonal)
ts_gas_ns<-(ts_gas_equiv[,1] - decomp_gas$seasonal)


```

## plot ACF and PACF to see the general pattern of two series
```{r}

# ACF and PACF of electricity data
par(mfrow=c(1,2))
Acf(ts_electricity[,2], lag.max=40, main="ACF Electricity")
Acf(ts_elec_ns, lag.max=40, main="ACF Non-Seasonal Electricity")

par(mfrow=c(1,2))
Pacf(ts_electricity[,2], lag.max=40, main="PACF Electricity")
Pacf(ts_elec_ns, lag.max=40, main="PACF Non-Seasonal Electricity")

# ACF and PACF of natural gas data
par(mfrow=c(1,2))
Acf(ts_gas_equiv[,1], lag.max=40, main="ACF Natural Gas")
Acf(ts_gas_ns, lag.max=40, main="ACF Non-Seasonal Gas")

par(mfrow=c(1,2))
Pacf(ts_gas_equiv[,1], lag.max=40, main="PACF Natural Gas")
Pacf(ts_gas_ns, lag.max=40, main="PACF Non-Seasonal Gas")
```

## Use seasonal Arima to model eletricity and natural gas
```{r}
#forecast Electricity SARIMA

arima.e.model<-auto.arima(window(ts_electricity[, 2], end=c(2022,1)))
arima.e.forecast<-forecast(arima.e.model, h=12)

autoplot(ts_electricity[, 2]) +
  autolayer(arima.e.forecast$mean, series = "SARIMA") +
  ylab("Price (cents/kWh)") +
  ggtitle("Electricity - SARIMA")
  # performance is not that good

#Gas SARIMA forecast
arima.gas.model<-auto.arima(window(ts_gas_equiv[, 1], end=c(2022,1)))
arima.gas.forecast<-forecast(arima.gas.model, h=12)

autoplot(ts_gas_equiv[, 1])+
  autolayer(arima.gas.forecast$mean, series = "SARIMA") +
  ylab("Price (cents/kWh)") +
  ggtitle("Natural Gas - SARIMA")
  # performance is not that good
```
### Examine seasonal Arima model's performance on electricity and NG data
```{r}

# model performance for electricity data
sarima_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(arima.e.forecast$mean)


# model performance for gas data
sarima_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(arima.gas.forecast$mean)

```



## Use Arima with Fourier terms to model
```{r}
# arima with fourier terms for eletricity
arima.e.four.forecast <- ts_electricity[, 2] %>% 
  window(end = c(2022, 1)) %>% 
  auto.arima(seasonal = FALSE, 
             xreg = fourier(window(ts_electricity[, 2], end = c(2022, 1)), 
                            K = 6)) %>% 
  forecast(xreg = fourier(window(ts_electricity[, 2], 
                                 start = c(2022,2)
                                 ),
                            K = 6),
           h = 12)

autoplot(arima.e.four.forecast$mean, series = "ARIMA with Fourier Terms") +
  autolayer(ts_electricity[, 2], series = "Historical Data") +
  ggtitle("Electricity - ARIMA with Fourier Terms") +
  ylab("Price (cents/kWh)")

# arima with fourier terms for gas
arima.gas.four.forecast <- ts_gas_equiv[, 1] %>% 
  window(end = c(2022, 1)) %>% 
  auto.arima(seasonal = FALSE,
             xreg = fourier(window(ts_gas_equiv[, 1], end = c(2022, 1)), 
                            K = 6)
             ) %>% 
  forecast(xreg = fourier(window(ts_gas_equiv[, 1], 
                                 start = c(2022, 2)
                                 ),
                          K = 6), 
           h = 12)

autoplot(arima.gas.four.forecast$mean, series = "ARIMA with Fourier Terms") +
  autolayer(ts_gas_equiv[, 1], series = "Historical Data") +
  ggtitle("Natural Gas - ARIMA with Fourier Terms") +
  ylab("Price (cents/kWh)")

```
### Examine Arima with fourier's performance on electricity and NG data
```{r}

# model performance for electricity data
arima_four_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(arima.e.four.forecast$mean)

# model performance for gas data
arima_four_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(arima.gas.four.forecast$mean)

```

## Use STL to model
```{r}
# STL for electricity
stl.e.forecast <- ts_electricity[, 2] %>% 
  window(end = c(2022, 1)) %>% 
  stlf(h = 12)

autoplot(stl.e.forecast$mean, series = "STL + ETS") +
  autolayer(ts_electricity[, 2], series = "Historical Data") +
  ggtitle("Electricity - STL + EST") +
  ylab("Price (cents/kWh)")

# STL for gas
stl.gas.forecast <- ts_gas_equiv[, 1] %>% 
  window(end = c(2022, 1)) %>% 
  stlf(h = 12)

autoplot(ts_gas_equiv[, 1], series = "Historical Data") +
  autolayer(stl.gas.forecast$mean, series = "STL + ETS") +
  ggtitle("Natural Gas - STL + EST") +
  ylab("Price (cents/kWh)")
```

### Examine STL's performance on electricity and NG data
```{r}

# model performance for electricity data
stl_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(stl.e.forecast$mean)

# model performance for gas data
stl_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(stl.gas.forecast$mean)

```

## Use ETS to model -- not finished
```{r}
# I think STL doesn't need fourier terms. So, I will get covariates dataframes without fourier terms
# because UKRWAR and temperature are the same across two data frames, we only need one
cov_train_nofour <- covariates_train_e[, 1:2]
colnames(cov_train_nofour) <- c("UKRWAR", "temperature")

cov_test_nofour <- covariates_full_e[, 1:2] %>% 
  window(start = c(2022, 2))
colnames(cov_test_nofour) <- c("UKRWAR", "temperature")
```

## Use Neural Network and fourier to model
```{r}
# neural network forecast for electricity data
nn.e.forecast <- ts_electricity[, 2] %>% 
  window(end = c(2022 ,1)) %>% 
  nnetar(p = 2, P = 1, # 2 and 1 are randomly decided. To find the optimal one, 
         # we need to run a couple of combinations to find the best one
         xreg = fourier(window(ts_electricity[, 2], 
                               end = c(2022, 1)
                               ),
                        K = 6)
         ) %>% 
  forecast(xreg = fourier(window(ts_electricity[, 2], 
                                 start = c(2022, 2)
                                 ),
                          K = 6),
           h = 12)

# neural network forecast for gas data
nn.gas.forecast <- ts_gas_equiv[, 1] %>% 
  window(end = c(2022 ,1)) %>% 
  nnetar(p = 2, P = 1,
         xreg = fourier(window(ts_gas_equiv[, 1], 
                               end = c(2022, 1)
                               ),
                        K = 6)
         ) %>% 
  forecast(xreg = fourier(window(ts_gas_equiv[, 1], 
                                 start = c(2022, 2)
                                 ),
                          K = 6),
           h = 12)


autoplot(nn.e.forecast$mean, series = "Neural Network") +
  autolayer(ts_electricity[, 2], series = "Historical Data") +
  ggtitle("Electricity - Neural Network") +
  ylab("Price (cents/kWh)")

autoplot(nn.gas.forecast$mean, series = "Neural Network") +
  autolayer(ts_gas_equiv[, 1], series = "Historical Data") +
  ggtitle("Natural Gas - Neural Network") +
  ylab("Price (cents/kWh)")
```

### neutral network model performance
```{r}

# neural network model performance for electricity data
nn_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(nn.e.forecast$mean)

# neural network model performance for gas data
nn_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(nn.gas.forecast$mean)
```

## Use Neural Network, temperature, UKRWAR, and fourier to model
```{r}
# neural network forecast for electricity data
nn.cov.e.forecast <- ts_electricity[, 2] %>% 
  window(end = c(2022 ,1)) %>% 
  nnetar(p = 2, P = 1,
         xreg = covariates_train_e) %>% 
  forecast(xreg = window(covariates_full_e, start = c(2022, 2)),
           h = 12)

# neural network forecast for gas data
nn.cov.gas.forecast <- ts_gas_equiv[, 1] %>% 
  window(end = c(2022 ,1)) %>% 
  nnetar(p = 2, P = 1,
         xreg = covariates_train_gas) %>% 
  forecast(xreg = window(covariates_full_gas, start = c(2022, 2)),
           h = 12)

autoplot(nn.cov.e.forecast$mean, series = "Neural Network") +
  autolayer(ts_electricity[, 2], series = "Historical Data") +
  ggtitle("Electricity - Neural Network") +
  ylab("Price (cents/kWh)")

autoplot(nn.cov.gas.forecast$mean, series = "Neural Network") +
  autolayer(ts_gas_equiv[, 1], series = "Historical Data") +
  ggtitle("Natural Gas - Neural Network") +
  ylab("Price (cents/kWh)")
```

### neutral network model, temperature, fourier, UKRWAR performance
```{r}

# neural network model performance for electricity data
nn_cov_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(nn.cov.e.forecast$mean)

# neural network model performance for gas data
nn_cov_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(nn.cov.gas.forecast$mean)
```

## use TBATS model 
```{r}
# TBATS for electricity
tbats.e.forecast <- ts_electricity[, 2] %>% 
  window(end = c(2022, 1)) %>% 
  tbats() %>% 
  forecast(h = 12)

# TBATS for natural gas
tbats.gas.forecast <- ts_gas_equiv[, 1] %>% 
  window(end = c(2022, 1)) %>% 
  tbats() %>% 
  forecast(h = 12)

```

### TBATS model performance
```{r}

# neural network model performance for electricity data
tbats_e_perf <- ts_electricity[, 2] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(tbats.e.forecast$mean)

# neural network model performance for gas data
tbats_gas_perf <- ts_gas_equiv[, 1] %>% 
  window(start = c(2022, 2)) %>% 
  accuracy(tbats.gas.forecast$mean)
```


## compare performance scores and generate tables for use
```{r}
# plot these together
ts_electricity[, 2] %>% 
  window(start = c(2016, 1)) %>% 
  autoplot(series = "Historical Series") +
  autolayer(nn.e.forecast$mean, series = "Neural Network") +
  autolayer(arima.e.forecast$mean, series = "SARIMA") +
  autolayer(arima.e.four.forecast$mean, series = "ARIMA with Fourier Terms") +
  autolayer(stl.e.forecast$mean, series = "STL + EST") +
  autolayer(tbats.e.forecast$mean, series = "TBATS") +
  autolayer(nn.cov.e.forecast$mean, series = "Neural Network with Covariates") +
  ylab("Price (cents/kWh)") +
  ggtitle("Comparision of Four Modeling Methods - Electricity") +
  theme_classic()

  
ts_gas_equiv[, 1] %>% 
  window(start = c(2016, 1)) %>% 
  autoplot(series = "Historical Series") +
  autolayer(nn.gas.forecast$mean, series = "Neural Network") +
  autolayer(arima.gas.forecast$mean, series = "SARIMA") +
  autolayer(arima.gas.four.forecast$mean, series = "ARIMA with Fourier Terms") +
  autolayer(stl.gas.forecast$mean, series = "STL + EST") +
  autolayer(tbats.gas.forecast$mean, series = "TBATS") +
  autolayer(nn.cov.gas.forecast$mean,series = "Neural Network with Covariates") +
  ylab("Price (cents/kWh)") +
  ggtitle("Comparision of Four Modeling Methods - Natural Gas") +
  theme_classic()

# scores for electricity
scores.e <- rbind(sarima_e_perf, 
                  arima_four_e_perf,
                  stl_e_perf,
                  nn_e_perf,
                  tbats_e_perf,
                  nn_cov_e_perf
                  ) %>% 
  as.data.frame()

  # rename rows
row.names(scores.e) <- c("SARIMA", 
                         "ARIMA with Fourier", 
                         "STL",
                         "Neural Network",
                         "TBATS",
                         "Neural Network with Covariates")

  # find the row index of the lowest RMSE
best.e.model <- scores.e$RMSE %>% 
  which.min() 

cat("The best model for electricity by RMSE is: ", 
    row.names(scores.e[best.e.model, ]))

# generate a visualized table to use in the report
kbl(scores.e,
    caption = "Forecast Accuracy for NC Residential Electricity Price",
    digits = array(5, ncol(scores.e))) %>% 
  kable_styling(full_width = FALSE, position = "center", 
                latex_options = "hold_position") %>% 
  kable_styling(latex_options = "striped", 
                stripe_index = best.e.model,
                stripe_color = "red")

# scores for gas
scores.gas <- rbind(sarima_gas_perf, 
                  arima_four_gas_perf,
                  stl_gas_perf,
                  nn_gas_perf,
                  tbats_gas_perf,
                  nn_cov_gas_perf) %>% 
  as.data.frame()

  # rename rows
row.names(scores.gas) <- c("SARIMA", 
                         "ARIMA with Fourier", 
                         "STL",
                         "Neural Network",
                         "TBATS",
                         "Neural Network with Covariates")

  # find the row index of the lowest RMSE
best.gas.model <- scores.gas$RMSE %>% 
  which.min() 

cat("The best model for natural gas by RMSE is: ", 
    row.names(scores.gas[best.gas.model, ]))

# generate a visualized table to use in the report
kbl(scores.gas,
    caption = "Forecast Accuracy for NC Residential Natural Gas Price",
    digits = array(5, ncol(scores.gas))) %>% 
  kable_styling(full_width = FALSE, position = "center", 
                latex_options = "hold_position") %>% 
  kable_styling(latex_options = "striped", 
                stripe_index = best.gas.model,
                stripe_color = "red")
```

## Use STL to model electricity and natural gas for the next 12 month
```{r}

e.forecast.2324 <- ts_electricity[, 2] %>% 
  stlf(h = 12)

gas.forecast.2324 <- ts_gas_equiv[, 1] %>% 
  stlf(h = 12)

e.forecast.2324$mean %>% 
  autoplot() +
  autolayer(e.forecast.2324$mean, series = "electricity forecast for 23-24") +
  autolayer(gas.forecast.2324$mean, series = "gas forecast for 23-24") +
  autolayer(window(ts_electricity[, 2], start = c(2017, 1)), 
            series = "historical data for electricity") +
   autolayer(window(ts_gas_equiv[, 1], start = c(2017, 1)), 
            series = "historical data for natural gas") +
  theme_classic() +
  ggtitle("Forecast of Electricity and Natural Gas Price for the Next 12 Months") +
  ylab("Price (cents/kWh)")
  


```

